**Generative AI Statement:** Used generative AI to help with parts of the website design, wording, and some code cleanup while building the Shiny app.

---
title: "ARMA vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ARMA vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width = 7.2, fig.height = 12)
library(armaVisualizer)
```

Before we start, we setup the variables needed for the demonstration.

```{r setup}

set.seed(42)
n <- 20
sigma <- 1.5
b <- c(0.2, 2) # b1, b0

# Data-generating process (DGP): ARMA(2,2)
p_arma22 <- c(0.60, -0.30)
q_arma22 <- c(0.45, -0.20)
```

This vignette uses one ARMA(2,2)-generated dataset, then compares five fitted models:

1. LR baseline (`p = 0, q = 0`)
2. ARMA(1,1) (`p = 1, q = 1`)
3. MA(2) (`p = 0, q = 2`)
4. AR(2) (`p = 2, q = 0`)
5. ARMA(2,2) (`p = 2, q = 2`)

## Generate ARMA(2,2) Errors and Dataset

```{r generate-errors}
e_arma22 <- generate_ARMA_errors(
  n = n,
  p = p_arma22,
  q = q_arma22,
  sigma = sigma
)

d_arma22 <- generate_ARMA_dataset(
  n = n,
  p = p_arma22,
  q = q_arma22,
  sigma = sigma,
  b = b
)
```

```{r plot-generated}
par(mfrow = c(2, 1))
plot_ARMA_series(e_arma22, series = "e")
plot_ARMA_series(d_arma22, series = "y")
```

## Fit LR, ARMA(1,1), MA(2), AR(2), and ARMA(2,2)

```{r fit-models}
data_arma <- d_arma22

fit_lr <- fit_ARMA(data_arma, p_order = 0, q_order = 0)

fit_arma11 <- fit_ARMA(data_arma, p_order = 1, q_order = 1)

fit_ma2 <- fit_ARMA(data_arma, p_order = 0, q_order = 2)

fit_ar2 <- fit_ARMA(data_arma, p_order = 2, q_order = 0)

fit_arma22 <- fit_ARMA(data_arma, p_order = 2, q_order = 2)

# Compare innovation variances
fit_lr$sigma2
fit_arma11$sigma2
fit_ma2$sigma2
fit_ar2$sigma2
fit_arma22$sigma2
```

## Compare Fitted Curves and Error Fits

```{r plot-model-fits}
par(mfrow = c(5, 2))

plot_ARMA_series(data_arma, series = "y", fit = fit_lr)
plot_ARMA_series(data_arma, series = "e", fit = fit_lr)

plot_ARMA_series(data_arma, series = "y", fit = fit_arma11)
plot_ARMA_series(data_arma, series = "e", fit = fit_arma11)

plot_ARMA_series(data_arma, series = "y", fit = fit_ma2)
plot_ARMA_series(data_arma, series = "e", fit = fit_ma2)

plot_ARMA_series(data_arma, series = "y", fit = fit_ar2)
plot_ARMA_series(data_arma, series = "e", fit = fit_ar2)

plot_ARMA_series(data_arma, series = "y", fit = fit_arma22)
plot_ARMA_series(data_arma, series = "e", fit = fit_arma22)
```

From these fit comparisons, the baseline LR model leaves clear serial structure in residuals.
ARMA(1,1), MA(2), and AR(2) improve fit to different degrees, while ARMA(2,2)
typically captures the DGP most closely.

## Bibliography

