---
title: "LR Generator Vignette"
author: "Eshan Hussain"
date: "February 25, 2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r source, include=FALSE}
source("../codebase/LR/generate_data.R")
```

This vignette describes how to use the `generate_data` function to create synthetic data.

## Generating data around a user defined LR equation

The function `generate_data()` adds random noise to y values and generates (x,y) pairs that follows a user-defined relation. It generates random independent `x` variables, computes dependent $y$ values using the equation $Y = XB + \epsilon$, and includes the noise term $\epsilon$. 

If `save` is TRUE, the function saves the data inside a csv file, else, it returns a dataframe containing an arbitrary amount of x values, y values, and the noise applied to the relation.

## Function Parameters
```markdown
- `num_x`: Number of x variables in the equation. For example, y = 2x + 1 has 1 x value.
- `b_vector`: Vector of coefficients including the intercept. For example, the b vector for y = 2x + 1 is `c(1, 2)`.
- `values`: Number of observations (y values) to generate.
- `sd`: Standard deviation of the error term.
- `save`: Boolean. If `TRUE`, saves output as a CSV file; if `FALSE`, returns a dataframe.
```
## Return Value
On success, the function either returns 0 if `save` is TRUE or a dataframe if `save` is FALSE. Note that the name of the csv file is `output.csv.`In the case where the length of `b_vector` does not equal to `num_x + 1`, the function returns `-1` on failure.


## Examples

### Example 1: Save to CSV
The example here generates 50 (x1, y) pairs. Each error follows a normal distribution with mean 0 and standard deviation 0.1. The (x1,y) pairs follow the y=3-2(x1) relation. It saves the result in `output.csv`.

```{r csv}
generate_data(num_x = 1, b_vector = c(3, -2), values = 50, sd = 0.1, save = TRUE)
```

```{r plot, echo=FALSE}
df <- read.csv("output.csv")

plot(df$x1, df$y, main = "y = 3 - 2x", xlab = "x", ylab = "y", col = "steelblue", pch = 16)

abline(a = 3, b = -2, col = "red", lwd = 2)

```

### Example 2: Return as Dataframe
We can create a new relation y=2x+1 and save it as a dataframe. We do this by setting `save` to FALSE. For this example, we will create 10 data points.

```{r dataframe}
generate_data(num_x = 1, b_vector = c(1, 2), values = 10, sd = 0.1, save = FALSE)
```

### Example 3: Multiple X Variables

You can generate data with multiple independent variables by increasing `num_x` and providing the appropriate number of coefficients. For instance, the example below generates data for this relation: y=1+2(x1)-(x2)

```{r example-multi}
df_multi <- generate_data(num_x = 2, b_vector = c(1, 2, -1), values = 100, sd = 0.5, save = FALSE)
head(df_multi)
```

### AI statement
I did not use generative AI to generate any code. I only used generative AI to learn new R functions and applied them through my understanding

### Citations
sanchez, A. nosedal. (n.d.). sta302-classical-SLR-assumptions. Lecture. Mississauga: University of Toronto.

sanchez, A. nosedal. (n.d.). sta302-estimation-and-GM-thm. Lecture. Mississauga: University of Toronto.

sanchez, A. nosedal. (n.d.). sta302-slr-in-matrix-terms. Lecture. Mississauga: University of Toronto.

