---
title: "Moving average vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Moving average vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../../STA380_Proposol/bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=9) 
```

Before we start, we setup the variables needed for the demonstration.

```{r setup}
source("../MA/moving_average.R")

set.seed(42)
n <- 20
sigma <- 10
burnin <- 50
b <- c(2, 2) # b1, b0
t <- seq(from = 1, to = n)

q_ma1 <- c(0.60)
q_ma2 <- c(0.60, -0.30)
```

This vignette uses one MA(2)-generated dataset, then compares three fitted models:

1. LR baseline (`q = 0`)
2. MA(1) (`q = 1`)
3. MA(2) (`q = 2`)

## Generate MA(2) Errors and Dataset

```{r generate-errors}
e_ma2 <- generate_MA_errors(n, q_ma2, sigma, burnin)
d_ma2 <- generate_MA_dataset(n, q_ma2, sigma, burnin, b)
```

```{r plot-errors}
par(mfrow = c(2, 1))
plot_MA_errors(e_ma2)
title(main = "Generated MA(2) errors over time")
plot_MA_dataset(d_ma2)
title(main = "Generated MA(2) dataset over time")
```

##  Fit LR, MA(1), and MA(2) on the Same MA(2) Dataset

```{r fit-models}
data_ma <- d_ma2
ma_fit_lr <- fit_MA(data_ma, q = numeric(0))
e_hat_lr <- as.numeric(residuals(ma_fit_lr))
y_hat_lr <- as.numeric(data_ma$y - e_hat_lr)

ma_fit_ma1 <- fit_MA(data_ma, q = q_ma1)
e_hat_ma1 <- as.numeric(residuals(ma_fit_ma1))
y_hat_ma1 <- as.numeric(data_ma$y - e_hat_ma1)

ma_fit_ma2 <- fit_MA(data_ma, q = q_ma2)
e_hat_ma2 <- as.numeric(residuals(ma_fit_ma2))
y_hat_ma2 <- as.numeric(data_ma$y - e_hat_ma2)

ma_fit_lr$sigma2
ma_fit_ma1$sigma2
ma_fit_ma2$sigma2

```

## Compare Fitted Curves and Error Fits

```{r plot-model-fits}
par(mfrow = c(3, 2))

plot_MA_dataset(d_ma2, y_hat_lr)
title(main = "MA(2) data with LR fit")
plot_MA_errors(d_ma2$e, e_hat_lr)
title(main = "True errors vs LR residuals")

plot_MA_dataset(d_ma2, y_hat_ma1)
title(main = "MA(2) data with MA(1) fit")
plot_MA_errors(d_ma2$e, e_hat_ma1)
title(main = "True errors vs MA(1) residuals")

plot_MA_dataset(d_ma2, y_hat_ma2)
title(main = "MA(2) data with MA(2) fit")
plot_MA_errors(d_ma2$e, e_hat_ma2)
title(main = "True errors vs MA(2) residuals")

```

From these fit comparisons, the LR residuals still retain clear MA(2) structure.
Moving to MA(1) improves the residual behavior but still leaves what looks like some MA dependence.
With MA(2), the residual pattern appears much closer to random noise.

## Bibliography